---
title: "L07 Scales, Axes & Legends"
subtitle: "Data Visualization (STAT 302)"
author: "Noah Holubow"
output:
  html_document:
    toc: true
    toc_float: true
    highlight: "tango"
---


```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, comment = FALSE)
```

## Overview

The goal of this lab is to explore ways to manage and manipulate scales, axes, and legends within `ggplot2`. 

## Datasets

We'll be using the `tech_stocks.rda`, `corruption.rda`, `cdc.txt`, and a few toy datasets.

<br>

```{r load-packages-data-seed}
# Load package(s)
library(tidyverse)

# Load datasets
load("data/tech_stocks.rda")
load("data/corruption.rda")

# Read in the cdc dataset
cdc <- read_delim(file = "data/cdc.txt", delim = "|") %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor")
  ))

# Set seed
set.seed(975312468)

# Selecting a random subset of size 100
cdc_small <- cdc %>% slice_sample(n = 100)

# Generating toy datasets for exercise 2
dat1 <- tibble(theta = c(0, 2 * pi))

dat2 <- tibble(
  theta = seq(0, 2 * pi, length.out = 100),
  obs = rnorm(100, sin(theta), 0.1),
  larger_than = ifelse(abs(obs) < abs(sin(theta)), "1", "0")
)
```

<br>

## Exercises

Complete the following exercises.

<br>

### Exercise 1

Using the `tech_stocks` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- Size of lines in legend is 1.3
- 0.75 and 0.85 will be useful
- Eliminated extra space in horizontal direction
- `key_glyph`

**Answer:**
```{r}
# plot
tech_stocks %>% 
  mutate(company = factor(company, levels = c("Facebook", "Alphabet", "Microsoft", "Apple"))) %>% # changing company order
  ggplot(aes(x = date, y = price_indexed)) + # basic plot
  geom_line(aes(color = company), key_glyph = "timeseries") + # plotting trend lines
  guides(color = guide_legend(override.aes = list(size = 1.3))) + # manually changing legend color
  labs(
    x = NULL,
    y = NULL
  ) + # labels
  scale_x_date(expand = c(0,0)) + # setting x-axis scale
  scale_y_continuous(
    position = "right", 
    breaks = scales::breaks_width(100), 
    labels = scales::label_dollar()) +
  theme_minimal() + # setting y-axis cale
  theme(
    legend.position = c(0.75, 0.85), 
    legend.title = element_blank()
    )
```


<br>

### Exercise 2



Using `corruption.rda` dataset, recreate the following graphic as precisely as possible.

*Hints:*
  
- Transparency is 0.6
- Only use 2015 data
- Point size is 3 in legend
- Color used is `grey40` & color palette `"Set1"`
- `"y ~ log(x)"` (method `"lm"`)
- Package `ggrepel`
- `box.padding` is 0.6
- Minimum segment length is 0
- seed is 9876

**Answer:**
```{r}
# countries
countries <- c("Iraq", "Argentina", "China", "Ghana", "Niger", "Chile", "Japan", "United States", "Singapore")

# label info
labels <- corruption %>% 
  filter(year == 2015,
         country %in% countries)

# plot
corruption %>% 
  filter(year == 2015) %>% # filtering out all years other than 2015
  ggplot(aes(x = cpi, y = hdi)) + # basic plot
  geom_point(aes(color = region), alpha = 0.6) + # plotting points
  geom_smooth(
    formula = "y ~ log(x)", 
    method = "lm", 
    color = "grey40", 
    se = FALSE) + # plotting line with log function
  labs(
    x = "Corruption Perceptions Index, 2015 \n (100 = least corrupt)",
    y = "Human Development Index, 2015 \n (1.0 = most developed)",
    title = "Corruption and human development (2015)"
  ) + # labels and title
  guides(color = guide_legend(override.aes = list(size = 3))) +  # manually changing legend
  ggrepel::geom_text_repel(
    data = labels, 
    aes(label = country), 
    box.padding = 0.6, 
    seed = 9876, 
    min.segment.length = 0) + # plotting country labels
  scale_color_brewer(palette = "Set1") + # manually setting color palette
  theme_minimal() +
  theme(
    legend.position = c(1,0), 
    legend.justification = c(1,0), 
    legend.title = element_blank()) # moving legend to bottom right corner
```


<br>
  
### Exercise 3

Use the toy datasets `dat1` and `dat2` to recreate the following graphic as precisely as possible.

*Hints:*

- Sizes used 2 (points) and 1.3 (line)
- Transparency 0.8
- Colors: `#56B4E9`, `darkgreen` (when `larger_than` is `1`), & `red` (when `larger_than` is `0`)

**Answer:**
```{r}
# plot
dat2 %>% 
  ggplot(aes(x = theta, y = obs)) + # basic plot
  geom_function(
    fun = sin, 
    color = "#56B4E9", 
    size = 1.3) + # plotting sine function
  geom_point(
    aes(color = larger_than), 
    size = 2, 
    alpha = 0.8) + # plotting points
  scale_color_manual(values = c("darkgreen", "red")) + # manually changing point colors
  labs(
    x =  "\u03B8",
    y = "sin(\u03B8)"
  ) + # labels
  theme_minimal() +
  theme(legend.position = "none")
```


<br>

### Exercise 4

Using `cdc_small`construct a scatterplot of `weight` by `height` with the following requirements: 

- Size of plotting characters should be 3.
- Color and shape should both identify `genhlth`.
- One legend for both color and shape.
- Legend title should be "General Health?" with a newline starting after general.
- Legend categories should be ordered from excellent (top) to poor (bottom) with each word in category capitalized in the legend.
- Legend should be placed in the lower right-hand corner of the plotting area. 
- Color should follow the `"Set1"` palette.
- Shape should have a solid triangle (17) for excellent, solid circle (19) for very good, an `x` (4) for poor, an hollow rotated square with an `x` in it (9) for fair, and a solid square (15) for good.
- `height` values should be limited between 55 and 80.
- `height` axis should display every 5th number between 55 and 80 and be appropriately labeled (i.e. `55 in`, `60 in`, ..., `80 in`). No axis title is necessary.
- `weight` values should be limited between 100 and 300. 
- `weight` axis should be on log base 10 scale, but still display weights in pounds starting at 100 and displaying every 25 pounds until 300. Must be appropriately labeled (i.e. `100 lbs`, `125 lbs`, ..., `300 lbs`). No axis title is necessary.
- Graph title should be `CDC BRFSS: Weight by Height`.
- Minimal theme.

**Answer:**
```{r}
# plot
cdc_small %>% 
  ggplot(aes(x = height, y = weight)) + # basic plot
  geom_point(
    aes(
      color = genhlth, 
      shape = genhlth), 
    size = 3) + # plotting shapes with colors
  scale_shape_manual(
    labels = c("Excellent", "Very Good", "Good", "Fair", "Poor"), 
    values = c(17, 19, 4, 9, 15)
    ) + # manually changing shapes
  scale_color_brewer(
    labels = c("Excellent", "Very Good", "Good", "Fair", "Poor"), 
    palette = "Set1"
    ) + # manually changing colors
  scale_x_continuous(
    labels = scales::label_number(prefix = "", suffix = " in", accuracy = 1), 
    breaks = scales::breaks_width(5), 
    limits = c(55,80)
    ) + # setting x-axis labels and intervals
  scale_y_log10(
    labels = scales::label_number(prefix = "", suffix = " lbs"), 
    breaks = scales::breaks_width(25), 
    limits = c(100,300)) + # setting y-axis labels and intervals
  labs(
    x = NULL, # no x-axis label
    y = NULL, # no y-axis label
    color =  "General \nHealth?", shape = "General \nHealth?", # legend label
    title = "CDC BRFSS: Weight by Height" # title
  )  +
  theme_minimal() +
  theme(
    legend.position = c(1,0), 
    legend.justification = c(1,0)
    ) # moving legend to bottom right corner
```



<br>
