---
title: "L06 Annotation & Positioning"
subtitle: "Data Visualization (STAT 302)"
author: "Noah Holubow"
output:
  html_document:
    toc: true
    toc_float: true
    highlight: "tango"
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, comment = FALSE, dpi = 300)
```

## Overview

The goal of this lab is to explore methods for annotating and positioning with `ggplot2` plots. This lab also utilizes `scale_*` to a greater degree which is part of our next reading. In fact, students may find going through/reading chapter [11 Colour scales and legends](https://ggplot2-book.org/scale-colour.html) useful.

## Datasets

We'll be using the `blue_jays.rda`, `titanic.rda`, `Aus_athletes.rda`, and `tech_stocks.rda` datasets.

<br>

```{r load-packages-data}
# Load package(s)
library(tidyverse)
library(patchwork)

# Load datasets
load("data/blue_jays.rda")
load("data/titanic.rda")
load("data/Aus_athletes.rda")
load("data/tech_stocks.rda")
```

<br>

## Exercises

Complete the following exercises.

<br>

### Exercise 1

Using `blue_jays.rda` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- Transparency is 0.8
- Point size 2
- Create a `label_info` dataset that is a subset of original data, just with the 2 birds to be labeled
- Shift label text horizontally by 0.5
- See 8.3 building custom annotations 
- Annotation size is 4
- Classic theme

**Answer:**
```{r}
# data for labels
label_info <- blue_jays %>% 
  filter(BirdID %in% c("1142-05914", "702-90567"))

# plot
blue_jays %>% 
  ggplot(aes(x = Mass, y = Head)) +
  geom_point(aes(color = KnownSex),
             show.legend = FALSE) +
  geom_text(data = label_info, aes(label = KnownSex, color = KnownSex),
            nudge_x= 0.5,
            show.legend = FALSE) +
  labs(
    x = "Body mass (g)",
    y = "Head length (mm)"
  ) +
  annotate(
    geom = "text", x = range(blue_jays$Mass)[1], y = range(blue_jays$Head)[2],
    label = "Head length versus body  mass for 123 blue jays", hjust = 0, vjust = 1, size = 4
  ) +
  theme_classic()
```


<br>

### Exercise 2

Using `tech_stocks` dataset, recreate the following graphics as precisely as possible.


<br>

#### Plot 1

*Hints:*

- Create a `label_info` dataset that is a subset of original data, just containing the last day's information for each of the 4 stocks
- serif font
- Annotation size is 4

**Answer:**
```{r}
# label info
label_info <- tech_stocks %>% filter(date == first(date))

# plot
tech_stocks %>% 
  ggplot(aes(x = date, y = price_indexed)) +
  geom_line(aes(color = company)) +
  labs(
    x = NULL,
    y = "Stock price, indexed"
  ) +
  geom_text(data = label_info, aes(label = company)) +
  theme_minimal() +
  theme(legend.position = "none") + 
  annotate(
    geom = "text", x = range(tech_stocks$date)[1], y = range(tech_stocks$price_indexed)[2],
    label = "Stock price over time for four major tech companies", hjust = 0, vjust = 1, size = 4,
    family = "serif"
  )
```

<br>

#### Plot 2

*Hints:*

- Package `ggrepel`
- Annotation size is 4
- `box.padding` is 0.6
- Minimum segment length is 0 
- Horizontal justification is to the right
- seed of 9876

**Answer:**
```{r}
# setting seed
set.seed(9876)

# mini stocks
label_info1 <- tech_stocks %>% filter(date == first(date))

# plot
tech_stocks %>% 
  ggplot(aes(x = date, y = price_indexed)) +
  geom_line(aes(color = company)) +
  labs(
    x = NULL,
    y = "Stock price, indexed"
  ) +
  ggrepel::geom_text_repel(data = label_info1, aes(label = company), box.padding = 0.6, min.segment.length = 0, hjust = 1) +
  theme_minimal() +
  theme(legend.position = "none") + 
  annotate(
    geom = "text", x = range(tech_stocks$date)[1], y = range(tech_stocks$price_indexed)[2],
    label = "Stock price over time for four major tech companies", hjust = 0, vjust = 1, size = 4,
    family = "serif"
  )
```


<br>

### Exercise 3

Using the `titanic.rda` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- Create new variable for that uses `died` and `survived` as levels/categories
- Hex colors: `#D55E00D0`, `#0072B2D0` (no alpha is being used)

**Answer:**
```{r}
# plot
# survived
titanic %>% 
  mutate(
    survived = factor(survived, labels = c("died", "survived")) 
  ) %>%
  ggplot(aes(x = sex)) +
  geom_bar(aes(fill = sex), show.legend = FALSE) +
  facet_grid(survived ~ class) +
  scale_fill_manual(values = c("#D55E00D0", "#0072B2D0")) +
  labs(
    x = NULL,
    y = NULL
  ) +
  theme_minimal()
```


<br>

### Exercise 4

Use the `athletes_dat` dataset --- extracted from `Aus_althetes.rda` --- to recreate the following graphic as precisely as possible. Create the graphic twice: once using `patchwork` and once using `cowplot`.

<br>

```{r include=FALSE}
# Get list of sports played by BOTH sexes
both_sports <- Aus_athletes %>%
  distinct(sex, sport) %>%
  count(sport) %>%
  filter(n == 2) %>%
  pull(sport)

# Process data
athletes_dat <- Aus_athletes %>%
  filter(sport %in% both_sports) %>%
  mutate(sport = case_when(
    sport == "track (400m)" ~ "track",
    sport == "track (sprint)" ~ "track",
    TRUE ~ sport
  ))
```

<br>

*Hints:*

- Build each plot separately
- Hex values for shading: `#D55E0040` and `#0072B240` (bottom plot),
`#D55E00D0` & `#0072B2D0` (for top two plots) --- no `alpha`
- Hex values for outline of boxplots: `#D55E00` and `#0072B2`
- Boxplots should be made narrower; 0.5
- Legend is in top-right corner of bottom plot
- Legend shading matches hex values for top two plots
- Bar plot lower limit 0, upper limit 95; shift bar labels by 5 units and top justify; label size is 5
- `rcc`: red blood cell count; `wcc`: white blood cell count
- Size 3 will be useful

<br>

#### Using `patchwork`

**Answer:**
```{r}
# colors
col <- c("#D55E00D0", "#0072B2D0")

# athlete count
count <- athletes_dat %>% 
  group_by(sex) %>% 
  summarise(count = n())
count

# bar graph
p1 <- athletes_dat %>% 
  ggplot(aes(x = sex)) +
  geom_bar(aes(fill = sex), show.legend = FALSE) +
  labs(
    x = NULL,
    sex = c("female, male"),
    y = "number"
  ) +
  scale_x_discrete(labels = c("female", "male")) +
  scale_y_continuous(limits = c(0,95)) +
  scale_fill_manual(values = c("#D55E00D0", "#0072B2D0")) +
  geom_text(data = count, aes(label = count, y = count), nudge_y = -5, size = 5, vjust = "top") +
  theme_minimal()

# scatterplot
p2 <- athletes_dat %>% 
  ggplot(aes(x = rcc, y = wcc)) +
  geom_point(aes(fill = sex), color = "white", size = 3, shape = 21, show.legend = FALSE) +
  scale_fill_manual(values = c("#D55E00D0", "#0072B2D0")) +
  labs(
    x = "RBC count",
    y = "WBC count"
  ) +
  theme_minimal()

# boxplot
p3 <- athletes_dat %>% 
  ggplot(aes(x = sport, y = pcBfat)) +
  geom_boxplot(aes(color = sex, fill = sex), width = 0.5) +
  scale_fill_manual(values = c("#D55E0040", "#0072B240"), name = NULL, labels = c("female", "male")) +
  scale_color_manual(values = c("#D55E00D0", "#0072B2D0"), name = NULL, labels = c("female", "male")) +
  guides(
    fill = guide_legend(ncol = 2, override.aes = list(fill = c("#D55E00D0", "#0072B2D0"), color = "transparent"))
  ) +
  theme_minimal() +
  theme(legend.position = c(1,1), legend.justification = c(1,1), legend.margin = margin(0)) +
  labs(
    x = "",
    y = "% body fat"
  )

# combining
(p1 + p2) / p3
  
```


<br>

#### Using `cowplot`

Use [`cowplot::plot_grid()`](https://cran.r-project.org/web/packages/cowplot/vignettes/plot_grid.html) to combine them

**Answer:**
```{r}
# plot
cowplot::plot_grid(cowplot::plot_grid(p1, p2), p3, nrow = 2)
```


<br>

### Exercise 5

Create the following graphic using `patchwork`.

*Hints:*

- Use plots created in exercise 4
- Useful values: 0, 0.45, 0.75, 1
- inset theme is classic

**Answer:**
```{r}
# bar graph
g1 <- athletes_dat %>% 
  ggplot(aes(x = rcc, y = wcc)) +
  geom_point(aes(fill = sex), color = "white", size = 3, shape = 21, show.legend = FALSE) +
  scale_fill_manual(values = c("#D55E00D0", "#0072B2D0")) +
  labs(
    x = "RBC count",
    y = "WBC count"
  ) +
  theme_minimal()

# scatterplot
g2 <- athletes_dat %>% 
  ggplot(aes(x = sex)) +
  geom_bar(aes(fill = sex), show.legend = FALSE) +
  labs(
    x = NULL,
    sex = c("female, male"),
    y = "number"
  ) +
  scale_x_discrete(labels = c("female", "male")) +
  scale_y_continuous(limits = c(0,95), expand = c(0,0)) +
  scale_fill_manual(values = c("#D55E00D0", "#0072B2D0")) +
  geom_text(data = count, aes(label = count, y = count), nudge_y = -5, size = 5, vjust = "top") +
  theme_classic()

g1 + inset_element(g2, left = 0.75, bottom = 0, right = 1, top = 0.45) +
  plot_annotation(tag_levels = 'A')
```


<br>
